Exploring the Validity Range of PLL Analytical Models

### 2025-05-30 CaseStudies/Example3_2_LimitsoftheTheorySoFar/Report/article/article.qmd
Almost finishing the report.

The loop filter transfer function, $F(s)$, neglecting the shunt capacitor, $C_2$ can be described as:

$$
F(s) = R+\frac{1}{sC_1}=\frac{sC_1R+1}{sC_1}
$$

Ainda ignorando o segundo capacitor, é possível escrever a função de transferência de malha aberta desse sistema apresentado na [Fig. @fig-PLL-integerN-blocks] como

@@

tal que $A_0$ é o ganho DC, $K_\theta=I/2\pi$ o ganho do charge pump, $K_\mathrm{VCO}$ o ganho do VCO e $N$ o divisor de frequência.

@@

$$
\begin{split}
H(s)&=\frac{\theta_o}{\theta_R}=\frac{G(s)}{1+G(s)}\implies\\
&=\frac{\frac{A_0K_\theta K_\mathrm{VCO}}{N}\left(R+\frac{1}{sC_1}\right)}{s+\frac{A_0K_\theta K_\mathrm{VCO}}{N}\left(R+\frac{1}{sC_1}\right)}
\end{split}
$$ {#eq-PLL-closed-loop-transfer-function-initial}

That can be rearanged as:

$$
H(s)=\frac{\frac{A_0K_\theta K_\mathrm{VCO}}{2\pi\cdot NC_1}\left(RC_1s+1\right)}{s^2+\frac{A_0K_\theta IK_\mathrm{VCO}}{2\pi\cdot N}Rs+\frac{A_0K_\theta IK_\mathrm{VCO}}{2\pi\cdot NC_1}}
$$ {#eq-PLL-closed-loop-transfer-function-modified}

ou, de maneira mais simples,

$$
H(s)=\frac{KF(s)}{s+KF(s)}
$$

Com o filtro simplificado, sem o secundo capacitor,

$$
H(s)=\frac{\frac{K}{C_1}\left(RC_1s+1\right)}{s^2+KRs+\frac{K}{C_1}}
$$

Using the parameters $\omega_n$ and $\zeta$, the closed-loop transfer function of the entire loop can be described as (neglecting the second capacitor, $C_2$, of the filter which would yield to a not wished third-order transfer function):

$$
H(s)=\frac{\omega_n^2\left(\frac{2\zeta}{\omega_n}s+1\right)}{s^2+2\zeta\omega_n s+\omega_n^2}
$$ {#eq-PLL-closed-loop-transfer-function-zeta-omega_n}

with

@@

For simplicity, the second capacitor can be obtained by $C_2=C_1/10$. Se esse segundo capacitor não fosse desconsiderado, os passos aqui apresentados seriam um pouco mais complicados. Entretanto, ainda it's possible to present the transfer function of the system in terms of all components values. This transfer function can be seen below (this equation uses the second capacitor, $C_2$, of the loop filter, adding a high-frequency pole to the system):

$$
\begin{split}
\frac{\theta_o}{\theta_R}&=\frac{K_\mathrm{VCO}K_\theta\left(1+sC_1R\right)}{s^2N\left(C_1+C_2\right)\left(1+sC_sR\right)+K_\mathrm{VCO}K_\theta\left(1+sC_1R\right)}
\end{split}
$$

where $C_s$ is the paralel between $C_1$ and $C_2$.

@@

Utilizando as expressões [Eq. @eq-PLL-closed-loop-transfer-function-zeta-omega_n] e [Eq. @eq-PLL-closed-loop-transfer-function-modified],

$$
\begin{split}
E(s)&=\frac{\frac{A_0K_\theta K_\mathrm{VCO}}{NC_1}\left(RC_1s+1\right)}{s^2+2\zeta\omega_ns+\omega_n^2}\cdot\frac{1}{1/N}\cdot\frac{1}{A_0K_\mathrm{VCO}/s}\cdot\frac{1}{F(s)}\cdot\frac{1}{K_\theta}\implies\\
&=\frac{\frac{A_0K_\theta K_\mathrm{VCO}}{NC_1}\left(RC_1s+1\right)}{s^2+2\zeta\omega_ns+\omega_n^2}\cdot N\cdot\frac{s}{A_0K_\mathrm{VCO}}\cdot\frac{sC_1}{RC_1s+1}\cdot\frac{1}{K_\theta}
\end{split}
$$

Resulting in:

$$
E(s)=\frac{s^2}{s^2+2\zeta\omega_ns+\omega_n^2}
$$ {#eq-phase-error-transfer-function}

@@

Multiplying the reference phase of [Eq. @eq-phase-of-frequency-step] to the [Eq. @eq-phase-error-transfer-function], results in:

$$
\begin{split}
\theta_e&=\frac{\Delta\omega}{s^2+2\zeta\omega_ns+\omega_n^2}
\end{split}
$$ {#eq-transfer-function-step-frequency-phase-error}

Getting the inverse Laplace transform of [Eq. @eq-transfer-function-step-frequency-phase-error], for $\zeta<1$, results in:

$$
\theta_e(t)=\Delta\omega\cdot\frac{1}{\omega_n\sqrt{1-\zeta^2}}e^{-\zeta\omega_n t}\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)
$$ {#eq-inverse-laplace-phase-error}

### Maximum Phase Error Instant

Given the value of $\omega_n$, the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

@@

### Maximum Normalized Phase Error

First of all, instead of using the [Eq. @eq-inverse-laplace-phase-error] and apply $t_m$, it's calculated the maximum normalized phase error, which is the same equation normalized by the input signal ($\Delta\omega$) and the natural frequency ($\omega_n$). So the maximum normalized phase error, $\theta_{e_\text{m,n}}$, is given by:

@@

### Maximum Allowed Frequency Step

This normalized value can be used in the following non-normalized equation for the phase error:

$$
\begin{split}
\theta_e(t)&=\frac{\Delta\omega}{\omega_n}\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt}
\end{split}
$$

The goal is to find the maximum input frequency step, $\Delta\omega_m$, such that the phase error does not exceed the maximum phase error that the PFD can withstand. Usually, this value is about $2\pi$. Using the normalized result:

@@

$$
V(s)=\frac{V_C}{\theta_R}=H(s)\cdot\frac{1}{1/N}\cdot\frac{1}{K_\mathrm{VCO}/s}
$$ {#eq-voltage-control-transfer-function-initial}

Utilizando as expressões [Eq. @eq-PLL-closed-loop-transfer-function-zeta-omega_n] e [Eq. @eq-PLL-closed-loop-transfer-function-modified],

$$
\begin{split}
V(s)&=\frac{\frac{A_0K_\theta K_\mathrm{VCO}}{NC_1}\left(RC_1s+1\right)}{s^2+2\zeta\omega_ns+\omega_n^2}\cdot\frac{1}{1/N}\cdot\frac{1}{K_\mathrm{VCO}/s}\implies\\
&=\frac{\frac{A_0K_\theta K_\mathrm{VCO}}{NC_1}\left(RC_1s+1\right)}{s^2+2\zeta\omega_ns+\omega_n^2}\cdot N\cdot\frac{s}{K_\mathrm{VCO}}
\end{split}
$$

Resulting in:

$$
V(s)=\frac{\frac{A_0K_\theta}{C_1}s\left(RC_1s+1\right)}{s^2+2\zeta\omega_ns+\omega_n^2}
$$ {#eq-voltage-control-transfer-function}

Multiplicando o step de frequência nessa função, é obtido

$$
\begin{split}
V_C(s)&=\frac{\Delta\omega\frac{A_0K_\theta}{C_1s}\left(RC_1s+1\right)}{s^2+2\zeta\omega_ns+\omega_n^2}
\end{split}
$$ {#eq-transfer-function-step-frequency-voltage-control}

@@

@@

@@

An integer-N Phase-Locked Loop (PLL) was developed with a division factor of $N=$ `{python} N`. The system was designed with a VCO gain of $K_\mathrm{VCO}=$ `{python} format(KVCO)`Hz/V, and a charge pump current of $I=$ `{python} format(I)`A. The phase detector gain is given by $K_\theta=I/2\pi$. Além disso, para que seja possível coincidir os valores calculados com as respostas obtidas na simulação, foi obtido um valor empírico para o ganho dc, $A_0=$ `{python} round(A0, 2)`.

@@

With the PLL values properly defined, the transfer function of this system is given by:

@@

@@

@@

@@

With the PLL values properly defined, the closed-loop transfer function of this system is given by:

@@

@@

@@

With the intention of conducting tests to verify the occurrence of cycle slipping, the minimum frequency of the reference signal is estimated. A ratio often quoted as being safe is 10:1 for $\omega_r/\omega_n$; thus, the minimum safe frequency for the reference signal is given by $f_r=$ `{python} format(f_r_safe)`Hz. The minimum allowed frequency is given by:

That is, the minimum reference frequency is $f_r\ge$ `{python} format(f_r_max)`Hz. Therefore, since a PLL operating with a reference frequency of $f_r=$ `{python} format(f_r)`Hz and an output frequency of `{python} format(f_o)`Hz is desired, the designed system must operate correctly.

>>>>

The loop filter transfer function, $F(s)$ can be described as:

$$
F(s) = \left(R+\frac{1}{sC_1}\right)\parallelsum\frac{1}{sC_2}
$$

É possível escrever a função de transferência de malha aberta desse sistema apresentado na [Fig. @fig-PLL-integerN-blocks] como

@@

tal que $A_0$ é o ganho DC (usualmente unitário), $K_\theta=I/2\pi$ o ganho do charge pump, $K_\mathrm{VCO}$ o ganho do VCO e $N$ o divisor de frequência.

@@

$$
H(s)=\frac{\theta_\mathrm{fb}}{\theta_R}=\frac{G(s)}{1+G(s)}
$$ {#eq-PLL-closed-loop-transfer-function-initial}

## Parameters and Components

Os parâmetros do sistema podem ser calculados simplificadamente como se seguem:

@@

For simplicity, the second capacitor can be obtained by $C_2=C_1/10$.

@@

@@

Getting the inverse Laplace transform of the respose of phase error transfer function for a step frequency, for $\zeta<1$, results in the following time-domain equation:

$$
\theta_e(t)=\frac{\Delta\omega}{\omega_n}\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt}
$$ {#eq-inverse-laplace-phase-error}

### Maximum Step Frequency

Analysing the maximum of the [Eq. @eq-inverse-laplace-phase-error], the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

@@

Sendo a máxima normalized phase error,

@@

Então

@@

$$
V(s)=\frac{V_C}{\theta_R}=H(s)\cdot\frac{1}{1/N}\cdot\frac{1}{K_\mathrm{VCO}}
$$ {#eq-voltage-control-transfer-function-initial}

@@

## Validade da teoria

Para que os modelos teóricos sejam válidos é necessário respeitar the minimum frequency of the reference signal is estimated. A ratio often quoted as being safe is 10:1 for $\omega_r/\omega_n$. However, the minimum allowed frequency is given by:

Se essa regra não for respeitada, os modelos lineares não têm a mesma garantia, necessitando que a amostragem do sinal também seja levada em consideração e o sistema tratado como um sistema discreto que realmente o é.

@@

## Project

@@

Para exemplificar os resultados obtidos até então, an integer-N Phase-Locked Loop (PLL) was developed with a division factor of $N=$ `{python} N`. The system was designed with a VCO gain of $K_\mathrm{VCO}=$ `{python} format(KVCO)`Hz/V, and a charge pump current of $I=$ `{python} format(I)`A. The phase detector gain is given by $K_\theta=I/2\pi$. Além disso, para que seja possível coincidir os valores calculados com as respostas obtidas na simulação, foi obtido um valor empírico para o ganho dc, $A_0=$ `{python} round(A0, 2)`.

@@

@@

A transfer function of the filter is given by:

@@

The open-loop transfer function of the system is given by:

@@

The closed-loop transfer function of the entire system is given by:

@@

@@

The transfer function of the phase error is given by:

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = E_sys.num[0][0] if isinstance(E_sys.num[0], (list, tuple)) else E_sys.num
den = E_sys.den[0][0] if isinstance(E_sys.den[0], (list, tuple)) else E_sys.den

# Exibe como equação no Jupyter
display(Markdown(f"$$E(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-phase-error-num-transfer-function}}"))
```

The transfer function of the voltage control is given by:

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = Vc_sys.num[0][0] if isinstance(Vc_sys.num[0], (list, tuple)) else Vc_sys.num
den = Vc_sys.den[0][0] if isinstance(Vc_sys.den[0], (list, tuple)) else Vc_sys.den

# Exibe como equação no Jupyter
display(Markdown(f"$$V_C(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-voltage-control-num-transfer-function}}"))
```

## Validade do modelo

The minimum safe frequency for the reference signal is given by $f_r=$ `{python} format(f_r_safe)`Hz. However, the minimum reference frequency is $f_r\ge$ `{python} format(f_r_max)`Hz. Therefore, since a PLL operating with a reference frequency of $f_r=$ `{python} format(f_r)`Hz and an output frequency of `{python} format(f_o)`Hz is desired, the designed system must operate correctly.

@@

```{python}
#| echo: false
#| label: fig-pll-step-response-alt
#| fig-cap: "Closed-loop phase step response of the designed PLL."
#| fig-align: center

@@
